### 2025-07-14 paper.tex
Split up into chapterwise tex files

\documentclass[12pt,openany,final]{book}
\usepackage[letterpaper,margin=1in,footskip=0.25in]{geometry}
\fontfamily{cmr}
\usepackage[final]{changes}
\usepackage{lipsum}
\usepackage[nottoc]{tocbibind}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{flafter}
\usepackage{floatrow}
\usepackage{float}
\floatstyle{plaintop}
\restylefloat{table}
%\usepackage{floatflt}

\usepackage{placeins}
\usepackage{siunitx}
\usepackage{graphicx}
\graphicspath{{figures/}}% input figure files
\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{appendix}
\usepackage{amsmath}  % usually already loaded
\newcommand{\norm}[1]{\left\lVert#1\right\rVert}
\usepackage{cases}
\usepackage{calc}
\usepackage{amssymb}
\usepackage{color}
\usepackage{rotating}
\usepackage{enumitem}
\usepackage{soul}
\usepackage{indentfirst}
\usepackage{hyphenat}
\usepackage{xspace}
\usepackage{subcaption}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{tabularx}
\newcolumntype{Y}{>{\centering\arraybackslash}X}
\usepackage{xcolor}
\usepackage{lineno}
\usepackage{setspace}
\usepackage{titlesec}
\usepackage{hyperref}
\usepackage[superscript]{cite}
\usepackage{lineno}
\titleformat
{\chapter} % command
[display] % shape
{\centering\sc\LARGE} % format
{\centering Chapter \thechapter} % label
{0.2ex} % sep
{
} % before-code
\titlespacing*{\chapter}{0pt}{0.5in}{0.25in} %{0.5in}{0.25in}
\titlespacing*{\section}{0pt}{0pt}{0pt}
\hbadness=99999
%\epstopdfDeclareGraphicsRule{.tga}{png}{.png}{
%    convert #1 \OutputFile
%}
\hypersetup{
    colorlinks,
    citecolor=black,
    filecolor=black,
    linkcolor=black,
    urlcolor=black
}
\usepackage[titles]{tocloft}
\setlength\cftbeforechapskip{\baselineskip}
\setlength\cftbeforefigskip{\baselineskip}
\setlength\cftbeforetabskip{\baselineskip}
\setlength\cftfigindent{0in}
\setlength\cfttabindent{0in}
\makeatletter
\setlength{\@fptop}{0pt}
\makeatother
\usepackage{url}
\urlstyle{same}

%\newcommand{\cmpersec}{$\frac{\text{cm}}{\text{s}}$~}
\newcommand{\cmpersec}{cm/s~}
\newcommand{\etal}{\textit{et al.}}
\newcommand{\PM}{$\pm$~}
\newcommand{\invnm}{$\text{\si{\nm}}^{-1}$~}
\newcommand{\nmeter}{\si{\nm}~}
\newcommand{\tbxmulticol}[3]
    {\multicolumn{#1}
                 {>{\centering\hsize=\dimexpr#1\hsize+#1\tabcolsep+\arrayrulewidth\relax}#2}
                 {#3}}
\newcommand{\ra}[1]{\renewcommand{\arraystretch}{#1}}
%\newcommand{\cmpersec}{$\frac{\text{cm}}{\text{s}}$~}
\renewcommand{\arraystretch}{1.8}
\newcommand{\sigmaij}{$\sigma_{ij}$}
\newcommand{\epsilonij}{$\epsilon_{ij}$}
\newcommand{\db}{$\text{D}_\text{B}$}
\newcommand{\dhh}{$\text{D}_\text{hh}$}
\newcommand{\dc}{$\text{2D}_\text{C}$}
\newcommand{\al}{$\text{A}_{\text{L}}$}
\newcommand{\vl}{$\text{V}_{\text{L}}$}
\newcommand{\vh}{$\text{V}_{\text{H}}$}
\newcommand{\vc}{$\text{V}_{\text{C}}$}
\newcommand{\sig}{$\sigma$}
\newcommand{\eps}{$\epsilon$}
\newcommand{\na}{Na\textsuperscript{+}}
\newcommand{\li}{Li\textsuperscript{+}}
\newcommand{\mg}{Mg\textsuperscript{2+}}
\newcommand{\cl}{Cl\textsuperscript{-}}
\newcommand{\mgcl}{MgCl\textsubscript{2}}
\newcommand{\nacl}{NaCl}
\newcommand{\licl}{LiCl}
\newcommand{\mbnbfix}{MB-NB-fix~}
\newcommand{\mbnbfixmicro}{Grotz \etal~} %CHANGED THIS
\newcommand{\nambnbfix}{Na\textsuperscript{+}--Saunders \etal}
\newcommand{\limbnbfix}{Li\textsuperscript{+}--Joung and Chetatham}
\newcommand{\mgmbnbfix}{Mg\textsuperscript{2+}--Li \etal} %CHANGED THIS
\newcommand{\mgmicro}{Mg\textsuperscript{2+}--Grotz \etal}
\newcommand{\mglbrules}{Mg\textsuperscript{2+}--Li \etal}
\renewcommand{\arraystretch}{1.8}
%\newcommand{\hl}{\added}
%\newcommand{\sout}{\deleted}
\newcommand{\TODO}{\hl{TODO}~}
\newcommand{\about}{$\sim$}
\newcommand{\aangstroms}{{\aa}ngstr{\"o}ms}
\newcommand{\aangstrom}{{\aa}ngstr{\"o}m}
\newenvironment{hle}{\color{red}}

\newcommand{\beginsupplemental}{%
    \setcounter{table}{0}
    \renewcommand{\thetable}{S\arabic{table}}%
    \setcounter{figure}{0}
    \renewcommand{\thefigure}{S\arabic{figure}}%
}
\renewcommand{\bibname}{References}
\titleformat{\chapter}
{\normalfont\normalsize\bfseries\centering}{\thechapter}{1em}{}
\titleformat{\section}
{\normalfont\normalsize\bfseries}{\thesection}{1em}{}
\titleformat{\subsection}
{\normalfont\normalsize\bfseries}{\thesubsection}{1em}{}
\titleformat{\subsubsection}
{\normalfont\normalsize\bfseries}{\thesubsubsection}{1em}{}
\titleformat{\paragraph}[runin]
{\normalfont\normalsize\bfseries}{\theparagraph}{1em}{}
\titleformat{\subparagraph}[runin]
{\normalfont\normalsize\bfseries}{\thesubparagraph}{1em}{}
%\cftsetpnumwidth{0.5in}
%\setlength{\cftbeforefigskip}{-\baselineskip}
%\setlength{\cftbeforetabskip}{-\baselineskip}
\cftsetpnumwidth{0.5in}
\cftsetrmarg{0.5in}
\renewcommand{\cftdot}{}
\renewcommand{\cftchapfont}{\normalsize}
\renewcommand{\cftchappagefont}{\normalsize}
%\setcounter{tocdepth}{3000}

\author{Matthew Saunders}
\begin{document}

\pagestyle{plain}
%\maketitle
\pagenumbering{roman}
\begin{titlepage}
\begin{centering}

~\\[1in]
Parameterizing Ion–Lipid Interactions from Local Clusters to Reproduce Bulk and Interfacial Properties in Lipid Bilayers
~\\[3\baselineskip]
by
~\\[3\baselineskip]
Matthew W. Saunders
~\\[4\baselineskip]
A thesis submitted in partial fufillment\\
of the requirements for the degree of\\
Doctorate of Philosophy in Biology \\ with a concentration in Cell \& Molecular Biology\\
Department of Molecular Biosciences\\
College of Arts and Sciences\\
University of South Florida
~\\[2\baselineskip]
Co-Major Professor: Sameer Varma, Ph.D.\\
Co-Major Professor: Sagar Pandit, Ph.D.\\
Jianjun Pan, Ph.D.\\
Libin Ye, Ph.D.\\
Stanley Stevens, Ph.D.
~\\[2\baselineskip]
Date of Approval:~\\
TBD
~\\[3\baselineskip]
Keywords: Molecular Dynamics, ???\\
~\\
Copyright \textsuperscript\textcopyright~2025, Matthew W. Saunders\\

\end{centering}
\end{titlepage}
\tableofcontents
\listoftables
\listoffigures
\bibliographystyle{elsarticle-num}
\doublespacing
\clearpage
\pagenumbering{arabic}
\chapter*{Abstract}
\addcontentsline{toc}{chapter}{Abstract}

\chapter[Introduction]{Introduction\footnote{Will I be reusing part of a publication here???}}
\input{ch1.tex}
\chapter[A high dimensional parameter search method to...]{A high dimensional parameter search method to determine force field mixing terms in molecular simulations\footnote{
Portions reprinted with permission from Matthew Saunders, Vered Wineman-Fisher, and Eric Jakobsson, \textit{High-Dimensional Parameter Search Method to Determine Force Field Mixing Terms in Molecular Simulations}, 
\textit{Langmuir}, American Chemical Society, March 1, 2022. 
\textcopyright{} 2022 American Chemical Society.
}}
\input{ch2.tex}
\chapter[Adsorption modes of \na, \li, and \mg to a model zwitterionic lipid bilayer]{Adsorption modes of \na, \li, and \mg to a model zwitterionic lipid bilayer\footnote{
    Portions reprinted with permission from Matthew Saunders, Abibat Adekeye-Olowofela, and Sabrina Downing, \textit{Adsorption Modes of Na\textsuperscript{+}, Li\textsuperscript{+}, and Mg\textsuperscript{2+} to a Model Zwitterionic Lipid Bilayer}, 
\textit{Langmuir}, American Chemical Society, December 1, 2024. 
\textcopyright{} 2024 American Chemical Society.
}}
\input{ch3.tex}
\chapter[Alteration of bilayer structure by adsorbed ions]{Alteration of lipid bilayer structure by adsorbed ions}
\input{ch4.tex}
\chapter{Conclusions}
\input{ch5.tex}
\bibliography{refs}
\end{document}