### 2025-07-14 ch2.tex
Split up into chapterwise tex files

Molecular dynamics (MD) force fields for lipids and ions are typically developed independently of one another. 
In simulations consisting of both lipids and ions, 
lipid-ion interaction energies are estimated using a predefined 
set of mixing rules for Lennard-Jones (LJ) interactions. 
This, however, does not guarantee their reliability. 
In fact, compared to the quantum mechanical reference data, 
Lorentz-Berthelot mixing 
rules substantially underestimate binding energies of \na~ ions with small 
molecule analogues of lipid headgroups, yielding errors on the order of 80 and 130 kJ/mol, 
respectively for methyl acetate and diethyl phosphate. Previously, errors 
associated with mixing force fields have been reduced using approaches like `NB-fix' 
in which LJ interactions are computed using explicit cross terms rather than 
those from mixing rules. Building on this idea, we derive explicit 
lipid-ion cross terms that also may implicitly include many-body 
cooperativity effects.
Additionally, to account for interdependency between cross terms, 
we optimize all cross terms simultaneously by 
performing high-dimensional searches using our ParOpt software. 
The cross terms we obtain reduce the 
errors due to mixing rules to below 10 kJ/mol. 
MD simulation of lipid bilayer conducted using these 
optimized cross terms resolve the structural discrepancies between our previous
simulations and small-angle X-ray and neutron scattering experiments. 
These results demonstrate that simulations of lipid bilayers with ions that are 
accurate up to structural data from scattering experiments 
can be performed without explicit polarization terms. 
However, it is worth noting that such NB-fix cross terms are
not based on any physical principle; a polarizable lipid model would be more 
realistic, and is still desired. 
Our approach is generic and can be applied to 
improve accuracies of simulations employing mixed force fields.

\section{Introduction}

Cellular membranes function as highly dynamic interfaces with many diverse components, 
including lipids, peptides, carbohydrates, and charged species like ionic salts.
Studies of these complex systems often benefit from computational methods, 
particularly molecular dynamics (MD) simulations~\cite{berkowitz:2019}. 
In our previous MD simulation studies, we characterized the effects of various 
monovalent and divalent ions on model 1-palmitoyl-2-oleoyl-sn-glycero-phosphatidylcholine (POPC) 
bilayers~\cite{pandit:2008:simulationtextbook,kruczek:2017,kruczek:2019,saunders:2019}. 
We reported that ions modify POPC bilayer structure with significant
effects on area per lipid and bilayer thickness. Similar results were also reported in MD simulations by
others~\cite{Bockmann:2003,cordomi:2008,gurtovenko:2008,Cordomi:2009,jurkiewicz:2012}.
Experiments characterizing bilayer structures in the presence of ions
have not been as numerous as simulation studies. However, experimental findings indicate that 
dissolved salts at physiological concentrations do not modify bilayer structure
significantly~\cite{pabst:2007,petrache:2006:swelling,uhrikova:2008}.
Specifically, Petrache \etal~ performed small angle X-ray scattering (SAXS) experiments on
multilamellar vesicles of 1,2-dilauroyl-sn-glycero-3-sn-glycero-phosphatidylcholine 
as well as other lipids in KCl and BrCl salt solutions, 
and reported that while small changes can be seen in the
X-ray scattering form-factor due to the 
salts, 
the fitted electron density profiles are essentially identical for
systems with and without 
salt~\cite{petrache:2006:swelling}. 
Similarly, Pabst \etal~found no significant change in bilayer structure 
for POPC bilayers in NaCl salt at or below 1 M concentration~\cite{pabst:2007}.
Furthermore, Uhrikova \etal~reported small structural changes using small angle neutron scattering
(SANS) experiments on 1,2-dipalmitoyl-sn-glycero-3-sn-glycero-phosphatidylcholine 
vesicles interacting with CaCl$_{2}$~\cite{uhrikova:2008}. 
Taken together, these results point to a general discrepancy between structural data 
from MD simulations and scattering experiments.

The reliability of MD simulations depends greatly on the force field (FF) parameters 
used for describing intra- and inter-molecular interactions. 
While FF parameters of lipids, including ours, are developed with great accuracy and care, 
we note that they are derived in the absence of ions. 
Similarly, ion parameters are also derived in the absence of lipids~\cite{joung:2008}. 
When simulations of bilayers are conducted in salt solution, 
ion-lipid interactions are computed using FF mixing rules. 
In our previous MD simulations of POPC bilayers in salt solutions, 
we employed our gromos43A1-S3 lipid FF parameters~\cite{chiu:2009} 
that were developed for use with SPC/E water to determine lipid-lipid and lipid-water interactions. 
Ion-ion and ion-water interactions were described using Joung and Chetham~\cite{joung:2008} parameters, 
also developed for use with SPC/E water. 
Lipid-ion interactions were estimated using Lorentz-Berthelot (LB) 
mixing rules for Lennard-Jones (LJ) components
, and there was a significant change in bilayer 
structure compared to that of the bilayer without salt despite the relatively small initial salt concentration of 200 mM.
Does this suggest that the discrepancy between our MD predictions and experiments 
is the result of the LB mixing rules?  
Note that none of the MD simulations of lipid-ion interactions discussed above
include explicit terms to describe electronic polarization. 
Errors in mixing rules may, therefore, 
emerge if the high electric fields of ions induce 
cooperativity effects in lipid groups 
differently from those in water. 
Quantum mechanical (QM) studies, in fact, suggest that many-body 
cooperativty effects, such as polarization depend \deleted{strongly} 
strongly on ion-coordinator chemistry~\cite{varma:2010,wineman:2019}. 
It has also been postulated that these effects, and specifically 
electronic 
polarization may play an
important role in determining the structure and dynamics of lipid
bilayers
-- especially when interacting with 
ions~\cite{vacha:2010,vorobyov:2010,melcr:2018,chen:2021,lee:2008:origin}.

Small deviations from LB rules have been shown to have a significant
    effect on the behavior of systems of particles\cite{boda:2008:effects}, and
    it is possible that a systematic tuning of these parameters could be used to correct 
    for artifacts in a simulation~\cite{baker:2010:accurate,yoo:2012:improved,fyta:2012:ionic,mamatkulov:2013:force,venable:2013,
savelyev:2014:balancing,li:2015:representation,savelyev:2015:competition,jing:2017:study,reif:2017,wineman:2019}.
Such a `Non-Bonded-fix' (NB-fix) strategy has been shown to effectively improve 
protein-ion 
, protein-nucleotide, and ion-membrane interactions 
while retaining the commonly used form of the LJ 6-12 potential~
\cite{baker:2010:accurate,yoo:2012:improved,fyta:2012:ionic,mamatkulov:2013:force,venable:2013,
savelyev:2014:balancing,li:2015:representation,savelyev:2015:competition,jing:2017:study,reif:2017,wineman:2019}. 
Building on this idea, here we propose a more 
general approach to optimize interaction cross terms
for use with the 6-12 potential, 
and also validate its prediction in condensed phase simulations.  
We expand on the NB-fix method by
(a) optimizing all ion-lipid LJ cross terms simultaneously, and 
(b) implicitly including many-body cooperativity effects.
We consider simultaneous optimization of all cross terms to be critical, 
because of their strong, interdependent correlation with the target results~\cite{fogarty:2014:paropt}.
This high-dimensional optimization is performed using our software tool
ParOpt~\cite{fogarty:2014:paropt,fogarty:2014:thesis}.
Many-body cooperativity effects have been shown to be a major contributor to ion 
binding~\cite{varma:2010}
. Thus, it is important to include them in lipid bilayer simulations where ions are known
to coordinate simultaneously with multiple ligands~\cite{kruczek:2019}.

We show that the cross terms we obtain from this approach substantially improve 
ion-lipid interaction energies over those obtained from LB mixing rules. 
MD simulation of a POPC bilayer
 in 200~mM NaCl initial solution 
conducted using these optimized cross terms also 
resolves the structural discrepancies between our 
previous MD simulations and small-angle X-ray and neutron scattering experiments 
at low salt concentrations.

\section{Methods}

The method proposed here is generic and can be applied to any pair of
interacting species that use cross terms, and ensures that we are
reproducing macroscopic results based on the most accurate
representation of the local inter-molecular interactions.  We chose
small molecular analogues of the important ion binding
sites in the polar region of phospholipid
molecules. These molecules were also used as building block molecules in development of our lipid
FF~\cite{kruczek:2017,chiu:2009}. Specifically, we selected
methyl--acetate (MeAc) to represent the ester group binding the
acyl--chain to the glycerol backbone, and diethyl--phosphate (DePh) to
represent the headgroup phosphate and surrounding carbons (See insert on figure~\ref{fig:energies}). 
The overall goal was to take the substitution energy of ions from
water to the selected molecules, along with the corresponding
geometries, all computed using a benchmarked quantum mechanical framework, 
and optimize the interaction cross terms to reproduce these target data 
within the Molecular Mechanics force-field.

Combined analysis of results from experiments and \emph{ab initio} molecular dynamics simulations
in the aqueous phase suggest that~\na~ions prefer to directly coordinate 
with $\sim 5-6$ water molecules~
\cite{varma:2006:coordination,mason:2006:neutron,galib:2017:revisiting,timko:2010:dissociation,smirnov:2020}.
However, when coordinating with MeAc molecules, steric hindrance restricts the 
number of binding partners to fewer than four coordinating molecules.  
Thus, we limited the size of our MeAc clusters to up to four molecules around an ion. 
DePh has resonant oxygens on each molecule that potentially act as two binding sites, so we
limited these clusters to up to two molecules around a \na. These were
compared to the clusters of \na surrounded by up to four water
molecules.
In this work, we forgo modifying terms for \cl, as we have found in our previous work 
that anions do not bind to the bilayer headgroup significantly, and remain solvated 
by water molecules~\cite{kruczek:2017}.

\subsection{Quantum Mechanical Calculations}

Target data for our parameter optimization consisted of energies and geometries 
computed using a benchmarked density functional theory as implemented in  
the FHI-Aims software package~\cite{fhiaims}. 
Geometry optimizations were performed on the \na--(Water)$_{n}$,
\na--(MeAc)$_{n}, n\leq4$, and \na--(DePh)$_m, m\leq2$ clusters.
These clusters were first optimized using the MM force field used in 
Kruczek \etal and Saunders \etal~\cite{kruczek:2017,saunders:2019}. 
MM optimized structures were then further
optimized using the PBE0 functional~\cite{perdew:1996:generalized,adamo:1999:toward} 
with self-consistent vdW corrections~\cite{tkatchenko:2009}. 
We used the \emph{really tight} basis sets included in the FHI--aims software. 
This functional and basis set combination 
has been shown to perform well compared to
experiment and high-level quantum methods for many different chemistries of ion-ligand clusters~\cite{wineman:2019,wineman:2020:transferable,wineman:2020:improved}. 
Optimizations were performed with a force 
maxima of $10^{-3}$ $\text{eV}/\text{\AA}$~, with total energies converged
to within $10^{-6}$~eV. 
We computed substitution energies of these clusters as:
\begin{equation}
    \begin{split}
        E^{n}_\text{MeAc}&=E_{\text{\na--(MeAc)}_n}-nE_\text{MeAc}-E_{\text{\na--(Water)}_n}+nE_\text{Water}\\ 
        E^{n}_\text{DePh}&=E_{\text{\na--(DePh)}_n}-nE_\text{DePh}-E_{\text{\na--(Water)}_{2n}}+2nE_\text{Water}\text{,}
    \end{split}
    \label{eq:substitution}
\end{equation}
where $n$ is the number of solvent molecules
(see tables~\ref{tab:qmbinding} and~\ref{tab:qmself} for all of the QM
data used for this computation).  These substitution energies and
corresponding configurations were used as targets for the parameter
optimization.

\begin{table}[h!]
    \caption[Total enegies of systems of small molecules...]{Total energies of systems of small molecules from QM calculations. 
These energies are computed by taking the total energy from the final step of geometry optimization 
on clusters of our selected small molecules around a single \na following
the procedure outlined in the methods section. We computed binding energies using these values.}
\label{tab:qmbinding}
\begin{tabularx}{\textwidth}{X|X|X|X}
    Cluster Size & Water (kJ/mol)& MeAc (kJ/mol) & DePh (kJ/mol)\\\hline
    1&$-9.048\times10^{5}$&$-1.130\times10^{6}$&$-2.528\times10^{6}$\\ \hline
    2&$-1.609\times10^{6}$&$-1.834\times10^{6}$&$-4.630\times10^{6}$\\\hline
    3&$-2.313\times10^{6}$&$-2.538\times10^{6}$& N/A \\\hline
    4&$-3.018\times10^{6}$&$-3.243\times10^{6}$& N/A \\\hline
\end{tabularx}
\end{table}
\begin{table}[h!]
    \caption[Self energies of isolated molecules]{Self 
    energies of isolated molecules. These are computed by performing geometry optimization on an isolated molecule following the procedure outlined in the methods.}
    \label{tab:qmself}
    {\footnotesize
\begin{tabularx}{\textwidth}{X|X|X|X|X}
               & Water         & \na           & MeAc          & DePh \\\hline
    Energy (kJ/mol)&-2.006E+05&-4.253E+05&-7.042E+05&-2.102E+06
\end{tabularx}
}
\end{table}

\subsection{Parameter Optimization}

Parameter optimization is performed using our ParOpt software package~
\cite{fogarty:2014:paropt,fogarty:2014:thesis}.
This software is available for download at
\texttt{https://csmlabfs1.cas.usf.edu/Sites}.
We utilized the Nelder-Meade method to perform a search to simultaneously optimize 
all \sigmaij~and \epsilonij~ cross terms of \na~ions with MeAc and DePh molecules. 
Specifically, there are seven atom types in these two small molecules (table~\ref{tab:paropt}), and so we optimized
14 cross terms for the 6-12 LJ potential.
Error was
determined by comparing the optimized
geometries and substitution energies of each new parameter set to the
reference data from QM.

Boundary constraints were imposed on \epsilonij~and \sigmaij~ to keep
the search space finite.  Table~\ref{tab:nmconstrain} shows all of the
constraints placed on the parameter search. 
Additionally, we
constrained the NA--OM\* \sigmaij~ to be smaller than the \sigmaij~ for
NA--P to avoid unphysical conformations of DePh.  Boundary constraints are enforced 
by reassigning \sigmaij~or \epsilonij~ values that violate the bound to the boundary value. 
Throughout optimization we monitored constraint
violations and ensured that we did not select a final parameter set
that is the result of a constraint violation.  High--dimensional optimizations of
this nature may not have a unique solution; thus, we performed 200
independent optimizations with random initial parameter values. We
compared the parameter sets that best improved the substitution energy
without significantly compromising the conformational geometries.
Table~\ref{tab:nmconstrain} shows all constraints used in this parameter search.
\begin{table}[H]
    \caption[Nelder-Meade constraints]{Nelder--Meade constraints. These values were used to constrain the parameter search space during the NM--optimization.}
    \label{tab:nmconstrain}
    {\tiny
    \begin{tabularx}{\textwidth}{X|X|X|X|X|X|X|X|X|X|}
	        &\tbxmulticol{2}{X|}{NA-CH3}&\tbxmulticol{2}{X|}{NA-CH2}&\tbxmulticol{2}{X|}{NA-CO*}&\tbxmulticol{2}{X|}{NA-OA,-OM*,-O*,-P}&Additional Constraints\\\hline
		&Min&Max&Min&Max&Min&Max&Min&Max&N/A\\\hline
	\sigmaij (nm)&0.2&0.5&0.2&0.5&0.2&0.5&0.2&0.5&$\sigma_{ij}^{\text{NA-OM*}}
        \leq \sigma_{ij}^{\text{NA-P}}$ \\\hline
	\epsilonij (kJ/mol) &0&0.79&0&0.81&0&0.83&0.05&7&N/A\\\hline
    \end{tabularx}
    }
\end{table}

Figure~\ref{fig:nmerror}a illustrates a representative
NM--trajectory that follows NM--error as a function of optimization step. In this case, the NM--error is
defined as an
equally--weighted combination of the mean absolute error of the
substitution energy and the distances between each atom in the cluster and the \na~ion. 
\begin{figure}[h!]
    \caption[Representative Nelder--Meade (NM) optimization run]{
    (a) Representative Nelder--Meade (NM) optimization run. Each point
    represents a move that the NM simplex can make while navigating the parameter space 
    (See Fogarty \etal~for a full description of the
    Nelder--Meade algorithm and available moves~\cite{fogarty:2014:paropt}). The insert illustrates the 
    RMSD between the simplex vertices.
    The optimization is considered converged when the simplex collapses, which is defined by an RMSD~$\leq 10^{-10}$.
    (b) Map of all \sigmaij~ and \epsilonij~ tested for interactions of \na~with non--carbon atom types 
    in the 200 optimizations performed to find our final optimized set of cross terms. A total of
291,870 combinations of parameters were tested, shown color-coded according to their NM error.}
    \label{fig:nmerror}
    %\includegraphics[width=\textwidth,trim=-3cm 0 0 -3cm]{7_NA_RERUN_CCONSTRAINED_EQ_3_ch1.eps}
    \includegraphics[width=\textwidth,trim=-3cm 0 0 -3cm]{figure_2_ch1.eps}
\end{figure}
Each NM--move used is illustrated
as a point on the error curve 
(see Fogarty \etal~for complete description of NM algorithm and
moves~\cite{fogarty:2014:paropt}).
The insert shows the root-mean squared distance (RMSD) between the simplex vertices at each step. 
As is typical with the NM method, error drops exponentially during the
initial steps, and slows down towards the end of the optimization
process. The termination condition for the optimization run is the
collapse of the NM--simplex (defined by the RMSD~$\leq 10^{-10}$).
Figure~\ref{fig:nmerror}b shows all of the 291,870 \sigmaij-\epsilonij~ pairs tested 
between \na~ions and the non--carbon atoms in the 200 independent optimization runs,
and provides a visual perspective of the sampled parameter space.
The parameter set that yielded the lowest error, as discussed in the results section, was
chosen to perform MD simulations of a POPC bilayer.

\subsection{Bilayer Construction}

We first constructed a monolayer of POPC lipids by placing 100 lipids on
a 10 nm by 10 nm grid, with excess space between the lipids to avoid overlaps in the lipid chains. 
Then we reflected this grid to create the second 
leaflet of the bilayer, resulting in a bilayer of 200 lipids.

Assuming a conservative estimate of one binding site per lipid, we
need at least 200 \na~ ions in bulk solvent at the beginning of the simulation
to avoid complete depletion of bulk ions after equilibration. 
In order to do this we constructed a system with
double the size of the solvent block used in our previous
works~\cite{kruczek:2017,kruczek:2019}.  This larger system was
constructed by adding 60,000 waters to the system on a 3-D grid with
    excess space between waters, and randomly
replacing water molecules with 216 \na and 216 \cl. 
This results in an initial concentration of 200~mM, similar to our previous simulations.
This process resulted in a simulation box with dimensions $9.75  \text{~nm}\times9.75  \text{~nm}\times59.84 \text{~nm}$.

We energy-minimized the simulation box using the steepest descent
algorithm with a force tolerance of 50 $\text{kJ mol}^{-1}\text{nm}^{-1}$. 
Neighbor searching was performed every 2 steps. 
The PME algorithm was used for electrostatic interactions with a cut-off of 1.6~nm. 
A reciprocal grid with a spacing of 0.12~nm$^{-1}$ was used with 6th order B--spline interpolation. 
A single cut-off of 1.6~nm was used for van der Waals interactions.

We then performed a 200~ps constant pressure simulation at 290~K to
ensure the system was relaxed enough for further annealing. 
The box dimensions at the end of this were $7.86\text{~nm}\times7.86\text{~nm}\times32.90$\text{~nm}.
Annealing was started at 400~K, and the system was cooled to the production
simulation temperature of 300~K in steps of 10~K. 
Each step was simulated for 150~ps, giving a total annealing time of 1.5~ns. 
The annealing process shrunk the box dimensions to $7.97\text{~nm}\times7.97\text{~nm}\times32.14\text{~nm}$.
This final structure 
was used as the starting point for production run.

\subsection{Molecular dynamics}

All molecular dynamics simulations were performed with the GROMACS software package, 
version 5.1.6~\cite{abraham:2015,pall:2014,van:2005,lindahl:2001,berendsen:1995}.
We have utilized the SPC/E model for all waters~\cite{spce}. Lipid interaction
terms are described using the parameters in the gromos43A1-S3
parameter set developed by our group in previous work~\cite{chiu:2009}. 
The system temperature was held constant at the production run 
temperature of 300K using the Nos\`e--Hoover thermostat
with a coupling constant of 0.5~ps~\cite{nose:1983}. 
Pressure coupling was performed using the Parrinello-Rahman semiisotropic barostat,
which held the system pressure constant at 1 atm with a coupling
constant of 1.5~ps~\cite{parrinello:1981}. 
The P--LINCS algorithm was used to constrain all bonds in the system to allow for a 4~fs
integration timestep~\cite{lincs}. 
Integration was carried out using the Verlet scheme, 
with neighbor-list updates taken on every other integration step. 
We used a cutoff of 16~\AA for short--range electrostatics. 
Beyond this cutoff, we have used the smooth particle--mesh 
Ewald summation method to describe electrostatics~\cite{essmann:1995}. 
LJ interactions were calculated with a cutoff of 16~\AA. 
For all systems described, we have simulated continuously for 0.7~$\mu$s.

Simulated trajectories were analyzed using a combination of GROMACS
built--in analysis tools and in--house software developed on the
GROMACS API.

\section{Results and Discussion}

\subsection{Optimized Cross-Terms}

The final optimized parameters are detailed in table~\ref{tab:paropt}
alongside the original parameters computed using LB rules.  
\begin{table}
    \caption[Force--field cross terms]{Force--field cross terms. Original terms, as used in the system simulated with LB rules were computed by applying Lorentz-Berthelot mixing rules to the LJ parameters of \na and each lipid component atom type. Optimized parameters are the result of the NM--optimization using ParOpt~\cite{fogarty:2014:paropt,fogarty:2014:thesis}. All constraints
    on the search space can be seen in table~\ref{tab:nmconstrain}.}
    \label{tab:paropt}
    {\footnotesize
    \begin{tabularx}{\textwidth}{X|X|X|X|X|}
              &\tbxmulticol{2}{X|}{Original}&\tbxmulticol{2}{X|}{Optimized}\\\hline
              &\sigmaij (nm) &\epsilonij (kJ/mol) &\sigmaij (nm) &\epsilonij (kJ/mol)\\\hline
        NA-CH3&0.295&1.100&0.235&0.700\\
        NA-CH2&0.312&0.772&0.237&0.809\\
        NA-OA &0.256&1.120&0.211&3.035\\
        NA-P  &0.277&1.900&0.301&0.483\\
        NA-OM*&0.252&1.221&0.211&1.445\\
        NA-CO*&0.335&0.362&0.315&0.758\\
        NA-O* &0.251&1.221&0.216&2.440\\
              \end{tabularx}
          }
\end{table}
We immediately note a general trend of an increase in the value of \epsilonij~ for the
non-carbon atom types. With our constraints on the carbon atoms, we have nudged the optimization
into gaining the binding energy by increasing the \epsilonij~ for the specifically electronegative atoms.
Values of \sigmaij~ have changed, but remained 
close to the original values in general, suggesting that the optimum 
distance to the minimum energy of the LJ potential is estimated well by LB rules. 
We can also see that no values of \sigmaij~ or \epsilonij~ violate the 
constraints described in table~\ref{tab:nmconstrain}. 
We examined substitution energies and corresponding conformational geometries by running energy
minimization of the QM--optimized structures using the final parameter set. 
These were then analyzed using the GROMACS built--in energy and distance tools. 
The substitution energies and the conformations for
this parameter set are shown in figure~\ref{fig:energies:ch1} 
and in figure~\ref{fig:distances:ch1}, respectively. 
\begin{figure}[h!]
    \caption[Substitution energies for \na~clusters computed as described
    in equation~\ref{eq:substitution}]{
Substitution energies for \na~clusters computed as described
in equation~\ref{eq:substitution}. 
In black we see the energies of systems computed using the standard 
mixing rules, in red we have the energies from benchmarked DFT, 
and in blue the optimized results. 
We see a significant
error with the standard LB mixing rules, which is substantially improved with our new optimized cross terms. 
The insert shows an diagram of POPC, and
the small molecules Methyl-Acetate (MeAc) and Diethyl-Phosphate (DePh) 
that were used to represent the major \na~interaction sites on the POPC molecule.}
    \label{fig:energies:ch1}
    %\includegraphics[width=\textwidth,trim=-3cm 0 0 0]{energies_combined_final_ch1.eps}
    \includegraphics[width=0.5\textwidth,trim=0cm 0 0 0]{figure_1_ch1.eps}
    %MOLECULE ON THE SIDE
\end{figure}
\begin{figure}[h!]
    \centering
    \caption[Distances from \na to each component atom in sample clusters.]{Distances from \na to each component atom in sample clusters. We compute the geometry of our sample clusters by computing the distance from
    the ion to each other atom in the system, shown per atom type.
    These distances are used in combination with the substitution energies in figure 2 to compute the
error for the NM optimization.}
    \label{fig:distances:ch1}
    %\includegraphics[width=0.8\textheight]{distances_combined_final_ch1.eps}
    \includegraphics[width=\textwidth]{figure_s1_ch1.eps}
\end{figure}
We can see that for MeAc we have substantially improved substitution
energies relative to those obtained from using LB mixing rules, 
which started with an discrepancy of around 30--80~kJ/mol. We have also
improved the relative substitution between the clusters of various sizes. 
The substitution energies for DePh have also improved by a similar magnitude. 
The conformational geometries are largely unchanged, 
with a general trend of the binding distance to OM\* shrinking on the order of 0.25~\AA~in DePh. 
This shrinkage is common when optimizing both energies and conformations
with the relatively small number of free parameters corresponding to
the LJ cross terms~\cite{wineman:2019}.

We also note that the substitution energies for both molecule types improve more in the
larger cluster sizes. Larger clusters are more relevant to the dense
environment in the lipid headgroup region of the bilayer, 
as few, if any, ions bind to a single lipid at a time~\cite{kruczek:2017}. 
Furthermore, the substitution energy profile for MeAc has become much closer to that of the QM profile. 
Thus, these new parameters substantially improve the energetic balance
between the lipid--ion, lipid--water and ion--water interactions.

The conformational geometries were mostly unchanged with the new
parameter set, as even the original parameters do a good job in reproducing the QM--configurations. 
The least precise cluster appears to be for 4 MeAc, 
where the original LB parameters poorly represent the symmetries exhibited in the QM data. 
Even with the improvement from our new parameters, 
we may be missing behavior from explicit polarization effects that
cannot be captured properly by a non--polarizable model~\cite{varma:2010}.

\subsection{Validation of Parameters}

In order to characterize our new parameter set in a bilayer, 
we generated a 700~ns simulation of a bilayer of POPC lipids in NaCl salt solution, 
and we compared the results against a similar system that we simulated using
LB rules in our previous works~\cite{kruczek:2017,kruczek:2019}. 
These older trajectories for systems both with salt and without will be referred to, 
respectively, as  LB and `without salt.'  
We simulated our system with optimized cross terms, 
hence forth will be referred to as the `optimized' system, 
long enough to equilibrate the number of bound ions (see figure~\ref{fig:ioncoodcount}). 
We will further characterize this ion binding in a subsequent section.

\subsubsection{Bilayer Structure}

Bilayer structural parameters can be seen in table~\ref{tab:struc}.
\begin{table}
    \caption[Bilayer structural parameters]{
        Bilayer structural parameters. \dhh~is the peak--to--peak distance from the electron density of the
        lipid bilayer, and is a measure of bilayer thickness. Bilayer thickness \db~and chain thickness \dc~are computed from
        number densities of the solvent and the lipid chains, respectively. $V_H$, and $V_C$ are the volumes of the headgroup and lipid chains computed
        using the method from Petrache \etal~\cite{petrache:1997}. $V_L$ is the sum of $V_H$ and $V_C$. 
        Rows 7-11 contain kinetic parameters for ion binding to membrane. These parameters come from fitting 
        the equation $ N_b(t)= \frac{K_{a}}{K_{a}+K_{d}} 
            N\left(1-\exp\left[-\left(K_{a}+K_{d}\right)\left(t-t_0\right)\right]\right)$ to the data for the number of ions
            bound to the lipid bilayer across the simulation time. 
            $A$ is the asymptotic number of ions bound to the lipid bilayer, and can be used as the expected number of ions
        that will bind to the system at equilibrium. $\tau$ is the characteristic timescale of the fitted function. $n_0$ is the number of ions bound at the beginning
        of the production run of the simulation. $K_D$ and $K_A$ are the computed binding association and dissociation constants, and $K_A/K_D$ is the binding
    rate constant.}
    \label{tab:struc}
    {\footnotesize
    \begin{tabularx}{\textwidth}{X|X|X|X}
                            & Without salt       & LB                                      & Optimized\\\hline
        $D_{HH}$~(\AA)      & 37.44 $\pm$ 1.07   & 40.18 $\pm$ 1.04                        & 37.64 $\pm$ 0.88 \\
        $D_B$~(\AA)         & 36.54 $\pm$ 0.47   & 40.90 $\pm$ 0.31                        & 39.36 $\pm$ 0.43\\
        $2D_C$~(\AA)        & 27.07 $\pm$ 0.34   & 30.33 $\pm$ 0.29                        & 28.97 $\pm$ 0.34     \\
        $V_H$~(\AA$^3$)     & 310.68 $\pm$ 1.14  & 316.13 $\pm$ 0.83                       & 314.81 $\pm$ 0.75          \\
        $V_C$~(\AA$^3$)     & 904.89 $\pm$ 1.28  & 891.79 $\pm$ 1.65                       & 896.50 $\pm$ 1.19         \\
        $V_L$~(\AA$^3$)     & 1215.57 $\pm$ 1.00 & 1207.92 $\pm$ 1.57                      & 1211.32 $\pm$ 1.21    \\
        $K_{A}$ (ns$^{-1}$) & N/A                & $7.12\times10^{-3}\pm8.18\times10^{-5}$ & $2.65\times10^{-3}\pm1.74\times10^{-5}$ \\
        $K_{D}$ (ns$^{-1}$) & N/A                & $3.20\times10^{-3}\pm4.75\times10^{-5}$ & $3.58\times10^{-3}\pm2.83\times10^{-5}$ \\
        $A$                 & N/A                & $74.51$                                 & $91.88$ \\
        $\tau$      (ns)    & N/A                & $96.73$                                 & $160.54$ \\
        $K_{A}/K_{D}$       & N/A                & $2.225$                                 & $0.74$ \\
        \end{tabularx}
        }
\end{table}
The phospholipid component volumes $V_H$ and $V_C$ (lines 1 and 2) are
computed following the procedure outlined by Petrache \etal~\cite{petrache:1997}. 
The lipid chains are identified as starting at the first carbon attached to the lipid chain carbonyl oxygen, including the oxygen.
The atom groups not part of the lipid
chains are partitioned into the headgroup volume. 
We take the number--density of these component groups along with that of the solvent, 
and use them to optimize the objective function:
\begin{equation}
    \label{eq:volumeobj}
    \Omega(v_i)=\sum^{\rho_s}_{z_j}(1-\sum^{N_{\text{Groups}}}_{i=1}{(\rho_i(z_j)v_i)^2})\text{,}
\end{equation}
In the equation above, $\rho_i(z_j)$ is the number density of the $i$ component in the
$z_j$ slice of the box and $v_i$ is the corresponding component volume. 
The component volumes are then multiplied by the corresponding
number of particles per molecule per group -- 32 for the chain
particles, and 20 for the headgroup. 
This gives us the total volume per molecule for each group. 
The total lipid volume $V_L$ (line 3 in table~\ref{tab:struc}) 
is taken to be the sum of these two values. 
These remain relatively similar in all three systems, 
as this value is intrinsic to the lipid molecule 
and should not change with the inclusion of ions.

Structural data are obtained for lipid bilayers via small angle X-ray and neutron scattering
experiments as a one--dimensional form--factor. Data are then
fitted to a continuous function to retrieve number and electron
densities for the various lipid components~\cite{nagle:2000,fogarty:2015}. 
Our simulations allow us direct access to the electron densities and number densities. 
The entries in table~\ref{tab:struc} are determined from these densities.

Figure~\ref{fig:eldens} shows the electron densities and corresponding bilayer form-factors. 
\begin{figure}[h!]
    \caption[Electron densities]{
    Electron densities of the simulated bilayers (a), and corresponding bilayer form--factors (b).
    Electron densities as obtained using the GROMACS density tool,
    centered at the minimum to define the bilayer center, and with the electron density of solvent subtracted.
    The simulated with optimized parameters appears to lack the large peak seen in the system simulated
    with LB rules, and appears more similar to the
bilayer structure of a bilayer simulated without salt. 
This is further reflected in the bilayer form--factor, 
computed by taking the cosine--transform of electron density. Experimental
SAXS results are for a POPC bilayer in pure solvent~\cite{fogarty:2015}. We see the first lobe
of the optimized system moves closer to the experimental results and the form--factor of a system without salt. This lines up with experimental results, that have shown small, if any, change
in the bilayer SAXS form--factor~\cite{pabst:2007,petrache:2006:swelling,uhrikova:2008}.}
    \label{fig:eldens}
    %\includegraphics[width=\textwidth,trim=-3cm 0 0 0]{ele_dens_ch1.eps}
    \includegraphics[width=0.5\textwidth,trim=-0cm 0 0 0]{figure_4_ch1.eps}
\end{figure}
Form--factors are computed by taking the cosine--transform of the symmetrized electron densities.   
We note that the simulations carried out using LB rules produced a
thicker bilayer and had different details at the peak region of the density. 
The new parameter set results in similar electron density to that of the system without salt.
This is similar to the results reported by Petrache
\etal~ and Pabst \etal, where for systems with less than 1~M NaCl, 
the differences in the electron densities were not
discernible~\cite{petrache:2006:swelling,pabst:2007}. 
These electron densities are used directly to measure the value of \dhh, 
defined as the peak--to--peak distance (see table~\ref{tab:struc} line 4). 
The new parameter set corresponds to a smaller \dhh, similar to the system without salt.

In addition to \dhh, different measures are used to assess the bilayer thickness 
that relies on the probability densities of different components of the system.
It can be shown that \db~(see table~\ref{tab:struc} line 5) computed by integrating one minus
the probability density of solvent and ions is equivalent to the computation of the 
Luzzati thickness of the total bilayer~\cite{fogarty:2015,chiu:2009}.
We define probability of finding a particular component in a slice of the box as,
\begin{equation}
    \label{eq:probability}
    P_{i}(z) = \frac{\rho_i(z)}{\sum^n_{j}\rho_{j}(z)}\text{,}
\end{equation}
where $\rho_{i}(z)$ are the number densities for the 
component particles ($i$) of the system as a function of the 
z-position of each slice of the box, and the summation
ranges over all components in the particular slice.
Thus, 
\begin{equation}
\text{\db}=\int_{\text{Box length}}\big(1-P_{\text{water}+\text{ions}}(z)\big)\;dz\text{.}
\end{equation}
In table~\ref{tab:struc} line 2, the \db~is larger for the systems
with ions, but the value obtained using our new parameter set is closer to that of
the bilayer simulated without salt.

We use a similar definition of probability density for \dc, 
computed from the probability distribution of the lipid chains. 
This component is defined by the hydrocarbon chains starting after the ester--linkage
on both the Sn1 and Sn2 terminal of the lipid backbone. 
This value (line 6 in table~\ref{tab:struc}) is increased in the system simulated
with LB rules over the system without salt, 
as we reported in our previous work. 
However, the new parameter set yields a value similar to the system without salt, 
which is consistent with the smaller overall thickness of the bilayer 
simulated with optimized cross terms.

The differences in bilayer thickness are closely related to the
packing of the lipid chains in the hydrophobic core of the bilayer. 
When the chains become more disordered, 
the bilayer thickness typically drops~\cite{nagle:2000}. 
Lipid chain ordering can be determined experimentally by performing NMR 
on specifically deuterated hydrocarbon chains. 
Since we lack hydrogen on our coarse--grained lipid chains, 
we cannot directly access the C--D ordering. 
Instead, we compute the chain order tensor $S_{\alpha\beta}$ defined as
\begin{equation}
    S_{\alpha\beta}=\frac{1}{2}\big\langle
    3\cos{\theta_\alpha}\cdot\cos{\theta_\beta}-\delta_{\alpha\beta}\big\rangle,
    \nonumber
\end{equation}
where the angles $\theta_\alpha$ and $\theta_\beta$ are the angles between the molecular
axis and the box z-direction. 
We then use this tensor to calculate the $S_{CD}$ as
\begin{equation}
    -S_{CD}^{\text{Saturated}}=\frac{2}{3}S_{xx}+\frac{1}{3}S_{yy}
\end{equation}
for saturated carbons~\cite{egberts:1988}, and as
\begin{equation}
-S_{CD}^{\text{Unsaturated}}=\frac{1}{4}S_{zz}+\frac{3}{4}S_{yy}\mp \frac{\sqrt{3}}{2}S_{yz}
\end{equation}
for unsaturated carbons~\cite{Douliez:1995}.  
These values are plotted per each carbon in the lipid chain in figure~\ref{fig:op:ch1}.
\begin{figure}[H]
    \caption[Acyl-chain order parameters]{ Lipid chain deuterium order parameters. $S_{CD}$s are computed for each carbon
        for the chains Sn1 (a) and Sn2 (b), starting at the second carbon in the chain. We see that the optmized system is still showing significant ordering in the lipid
    chains as a result of ion binding; however, the ordering is less pronounced than in the system simulated with LB rules, 
    and is closer to that of the simulation without salt. This
result corresponds with the smaller bilayer thickness in the optimized system.
}
    \label{fig:op:ch1}
    %\includegraphics[width=\textwidth,trim=-3cm 0 0 0]{chainorder_ch1.eps}
%    \includegraphics[width=\textwidth,trim=-3cm 0 0 0]{figure_4_ch1.eps}
    \includegraphics[width=0.5\textwidth,trim=-0cm 0 0 0]{figure_s2_ch1.eps}
\end{figure}
As reported in our previous simulations, 
the addition of salt has an ordering effect on the lipid chains. 
This effect is also seen in our new parameter set; however, the ordering is less pronounced, 
which is consistent with the notion that the bilayer structure is not significantly altered 
at physiological salt concentration~\cite{pabst:2007,petrache:2006:swelling}.

While this result indicates a structure more consistent with experimental results, 
the detailed structure of a lipid bilayer is a result of the
delicate balance between ion--lipid, lipid--water, and ion--water interactions.
In order to fully understand how our new parameter set has altered the
overall bilayer structure, we next characterize the specific interactions
between these moieties.

\subsection{Membrane-Salt Interactions}

Both ions and solvent compete for the binding sites on the lipid headgroup. 
As seen in figure~\ref{fig:energies}, the new cross terms
produce a relatively stronger interaction between \na and lipid
headgroup components compared to that of the LB rules. 
Thus, there is potentially a reduction in the available binding sites for the solvent. 
To examine how the new cross terms have altered ion interactions with lipids in the bilayer, 
we first characterize the dynamics of ion binding to the lipid bilayer.

We define ion binding to the lipid bilayer when half or fewer
of its first shell coordinators are not waters. 
In order to compute the equilibrium binding constant, 
we must determine the equilibrium number of bound ions to the lipid surface. 
Figure~\ref{fig:ioncoodcount} shows the number of bound ions as a
function of time over the entire duration of the simulation. 
\begin{figure}[h!]
    \caption[Number of ions bound to the bilayer vs time]{
Number of ions bound to the lipid bilayer as a function of simulation time. 
The exponential fits to this data are also shown. These fits are used to compute the asymptotic
number of ions bound as well as binding rate constants.
`Total' refers to the total number of ions in each simulation box.
A membrane bound ion is defined as having half or fewer of its first coordination shell
occupied by water molecules.}
    \label{fig:ioncoodcount}
    %\includegraphics[width=\textwidth,trim=-3cm 0 0 0]{ion_cood_count_ch1.eps}
    \includegraphics[width=0.5\textwidth,trim=-0cm 0 0 0]{figure_3_ch1.eps}
    %RENAME n(t) to fit
\end{figure}
We note that even after 700~ns of simulation time, the number of bound ions
are not fully equilibrated. 
Thus, we use first--order reaction kinetics to
estimate the asymptotic number of bound ions.  
The first--order reaction kinetics are modeled as a differential equation:
\begin{equation}
    \frac{dN_b}{dt}=K_a \left(N-N_b\right)- K_d N_b,
\end{equation}
where $N_b$ are the number of bound ions, and $K_a,K_d$ are the association and dissociation
time constants, respectively.
The solution of this differential equation is:
\begin{equation}
    N_b(t)= \frac{K_{a}}{K_{a}+K_{d}} N
    \left(1-\exp\left[-\left(K_{a}+K_{d}\right)\left(t-t_0\right)\right]\right).
    \label{eq:ioncoodnumfit}
\end{equation}
This solution is fit to the data in figure~\ref{fig:ioncoodcount}, 
and the resulting fit is also plotted. 
The fitting parameters are listed in table~\ref{tab:struc}.
The first--order reaction kinetic model fits reasonably well to the 
data from both the systems, except in the beginning of
the simulation where the effect of the annealing process is more pronounced; 
however, we are only interested in the asymptotic
behavior of the fit as this is representative of the equilibrium state of the system. 
The asymptotic number of bound ions as $t\to\infty$, $A=\frac{K_{a}}{K_{a}+K_{d}} N$ 
(table~\ref{tab:struc} row 9), is larger in the system simulated with optimized terms.
We also report the timescale of ion binding $\tau=\frac{1}{\left(K_{a}+K_{d}\right)}$
for both systems (table~\ref{tab:struc} row 10). 
The timescale of binding in the system using optimized  cross terms is longer, 
and suggests that this system would need more time to equilibrate than the system simulated with LB rules.
Finally, we report the value of $\frac{K_a}{K_d}$ 
(table~\ref{tab:struc} row 11), which
we observe is much smaller with the new parameter set than compared to that
of the system simulated with LB rules.

To examine how specific interactions between ions and lipids are modified by the new parameters, 
we tracked the binding partners of ions across the box over the last 150~ns of simulation time. 
Moieties are considered to be bound to an ion if they are within a distance of 3.3 \AA~from the \na~ion. 
Several electronegative groups in the simulation can potentially bind to the \na~ion. 
We compute the number of these potential binding partners within the first shell of each \na~ion
across the simulation box. 
Ions are then sorted according to their box z-positions, and then the data are averaged over the last 150~ns. 
This is plotted in figure~\ref{fig:cood}. 
\begin{figure}[h!]
    \caption[\na inner shell coordination]{
Chemistry of \na~inner shell coordination as a function of distance from bilayer
center. Compared to the system simulated with LB rules (a), the system simulated with optimized cross terms (b)
        yields a lower \na~total coordination number within the headgroup region of the bilayer. 
    This drop in coordination appears to be due to a greater dehydration of the ions in this system.}
    \label{fig:cood}
    %\includegraphics[width=\textwidth,trim=-3cm 0 0 0]{cood_ch1.eps}
    \includegraphics[width=0.5\textwidth,trim=-0cm 0 0 0]{figure_5_ch1.eps}
\end{figure}
We note first that the total number of solvating oxygens of ions within the bilayer headgroup
region with the optimized parameter set has dropped by $\sim 1$ when
compared to ions in similar locations in the system simulated using LB rules. 
This is not surprising, given the dependence of ion coordination
preferences on the local environment~\cite{varma:2008:JACS}. 
The binding to other lipid oxygens has not been altered much by the new parameter set; 
however, we do note that water within the headgroup region does not appear to be strongly associated with ions.

\subsection{Water Structure and Dynamics}

To further characterize the dehydration of ions in the new simulated system, 
we look to the lipid-- and ion--water interactions.
Figure~\ref{fig:waterdens} shows the number density of water as a
function of distance from the bilayer center for each of our simulated
systems, with the \dc~and \db~illustrated as dotted lines.  
\begin{figure}[h!]
    \caption[Water density at the bilayer interface]{
    Water density at the bilayer interface. We illustrate the regions regions $B_{-1}$, $B_+$, $B_{-2}$ and $Bulk$ for each system with dotted lines.
    We see that the optimized cross terms yield a greater density 
    of solvent in the $B_{+}$ and $B_{-1}$ regions over the system simulated with LB rules. 
We also see the density in these regions of the system optimized
with optimized cross terms is more similar to that of the system without salt.}
    \label{fig:waterdens}
    %\includegraphics[width=\textwidth,trim=-3cm 0 0 0]{dens_water_ch1.eps}
    \includegraphics[width=0.5\textwidth,trim=-0cm 0 0 0]{figure_6_ch1.eps}
\end{figure}
We see that our new parameter set produces a bilayer
interface that has more solvent inside the headgroup region, 
between $10-25$~\AA~from the bilayer center. 
This density is more similar to that of the system simulated without salt. 
This suggests that the dehydration of ions in the system simulated with optimized parameters
does not correspond to a dehydration of the lipid bilayer.

Next, we characterize the orientational structure of the water.
Figure~\ref{fig:waterorder} examines the water order parameter across the simulation box. 
\begin{figure}[h!]
    \caption[Water orientational order parameters]{
Water orientational order parameters $P_1$ (a) and $P_2$ (b), 
and the product of the water number density and $P_2$ (c). We see in $P_1$ and $P_2$ less
    ordering in the waters in the optimized system, suggesting that waters may be less strongly interacting with ions or lipid components. We denote the four regions of the lipid bilayer based on the shape of the $P_2$ data as dotted lines
    in (b)~\cite{saunders:2019}. We have not included these regions
 for the system without salt, as the $P_2$ data does not include the same details as the systems with salt. The integral of (c) is related to the
 quadrupolar splitting constant $\Delta \nu$ found in deuterium NMR experiments. This also gives a closer look at how solvent is ordered in the headgroup while
 accounting for the amount of solvent in the region. We see that optimized cross terms result in a significant drop in the area under the curve, which is much closer
 to the shape of the data from the system without salt. 
 The regions $B_{-2}$ and $Bulk$ are not within the bilayer
headgroup, and are expected to be less affected by the new parameter set.}
    \label{fig:waterorder}
    %\includegraphics[height=0.9\textheight,trim=-3cm -3cm 0 0]{h2order_ch1.eps}
    \includegraphics[width=0.5\textwidth,trim=-0cm -0cm 0 0]{figure_7_ch1.eps}
    %SAME line thickness as others
\end{figure}
We identify perturbed water structure by examining first ($P_1$) and second ($P_2$) 
orientational order parameters for the
OW$\rightarrow$HW1 bond of water with respect to the z-axis of the simulation box ($\beta$). 
These order parameters are defined using the first and second 
Legendre polynomials with respect to the angle $\beta$,
\begin{equation}
    \begin{split}
    P_1&=\langle \cos\left(\beta\right) \rangle
    \\ P_2&=\frac{1}{2}\bigg\langle\left(3\cos^2\left(\beta\right)-2\right)\bigg\rangle
    \text{,} 
\end{split}
\end{equation}
where average is over all the waters in a particular volume slice of the box 
and then over simulation time.  
We plot these values as a function of distance from the bilayer center. 
$P_1$ denotes dipolar ordering of the bond vector and the bilayer normal direction, 
with a positive value indicating an average outward orientation 
and a zero value corresponding to an average perpendicular orientation to the
bilayer normal or a uniformly random orientation. 
We observe a similar pattern of ordering across the box in all systems; 
however, we see an overall reduction in ordering with our new parameter set when
compared to both the LB and the no--salt system. 
We also see the inner minimum of the order parameter moved further into the bilayer
when compared to LB, 
which is consistent with the larger quantity of water 
in this region that we observe in the water densities.

Following the protocol established in our previous work~\cite{saunders:2019}, 
we identify three regions within the bilayer interface, $B_{-1}, B_{+}, B_{-2}$. 
The $B_{-1}$ region is defined as the region of negative ordering starting at the bilayer center, and ending 
    when the order parameter values cross zero at the start of the $B_{+}$ 
    region. The $B_{+}$ region starts at the end of the $B_{-1}$ region, and 
    is the area of positive ordering, ending where the order parameter crosses
    zero again. The $B_{-2}$ region starts at the end of the $B_{+}$, and extends out to where the 
    second order parameter goes to zero. This was found by fitting an exponential function to this region 
    and taking the \replaced{scale parameter}{timescale} from that fit as the boundary with bulk solvent.
We find that water is significantly less perturbed by the bilayer with our new parameter set.  
We have also computed $P_2\cdot\rho_{\text{Water}}$, shown in shown in figure~\ref{fig:waterorder}(c). 
This value relates the amount of water in each region of the box and 
the overall ordering in the region. 
We still see significantly less ordering with the new parameter set, 
and even with the larger number of waters in the bilayer headgroup. 
The integral of this curve is related to the quadrupole splitting $\Delta \nu$ observed in
in deuterium NMR experiments~\cite{aaman:2003,kruczek:2017}.

This suggests that while there is more solvent in the interface, it is perhaps not
associated with either \na or lipids, and may remain less structured than
in the system simulated with LB rules.  
This can be further ascertained by the diffusion coefficients of waters 
in each of the regions defined by $P_2$. 
We compute the mean square displacement (MSD) for water oxygens
in each region by first tracking which waters remain in the region. 
Any waters that leave the region are removed from the MSD calculation. 
We chose a duration of 100~ps to track the MSD in order
to have a sufficiently long time for the MSD to become linear, 
while still maintaining a statistically significant number of waters in the slice. 
A line is fit to the middle 80\% of the MSD, and the fitted
slope is used to calculate the diffusion coefficient following
Einstiein's relation for 2D diffusion
\begin{equation}
\lim_{t \to \infty}\frac{\big\langle (r(t) - r(0))^2 \big\rangle}{(t-t_0)} = 6D. 
\end{equation}
These values can be seen in table~\ref{tab:diff}. 
\begin{table}[h!]
    \caption[Diffusion coefficients of water]{
Diffusion coefficients of water in different regions of the lipid bilayer, 
defined by the shape of the second orientational order parameter of water molecules in the box. 
These regions are defined by the shape of the distribution of the second 
    orientational order parameter across the simulation box. $B_{-1}$ is the region 
    of negative ordering starting at the bilayer center, and ending 
    when the order parameter values cross zero. 
    $B_{+}$ starts at the end of the $B_{-1}$, and is the region of positive ordering ending where the order parameter 
    becomes negative. This starts the $B_{-2}$ of negative ordering, extending out to where the 
second order parameter goes to zero, where we have $Bulk$ solvent.
We see that the optimized parameters result in slightly 
    increased diffusion in the solvent, which correlates with the reduced
    ordering of the water dipoles and quadrupoles in the
system.}
    \label{tab:diff}
    {\footnotesize
    \begin{tabularx}{\textwidth}{X|X|X|}
              &LB ($\times 10^{-10} m^2/s$)&Optimized ($\times 10^{-10} m^2/s$)\\\hline
        $B_{-1}$ &1.11 $\pm$ 1.10& 1.88 $\pm$ 2.41   \\
        $B_+$    &4.23 $\pm$ 1.14& 6.11 $\pm$ 2.83  \\
        $B_{-2}$ &18.11 $\pm$ 4.23&21.29 $\pm$ 4.12 \\
        $Bulk$   &27.32 $\pm$ 1.15 &27.25 $\pm$ 1.36 \\
              \end{tabularx}
          }
\end{table}
We note that the water in the headgroup region, corresponding to $B_{-1}$ and $B_+$, 
diffuses slightly faster with the new parameter set, 
indicating more mobile water in these regions. 
However, the computed diffusion coefficients are within the error bars that of the system simulated with LB rules. 
Diffusion in the $B_{-2}$ and $Bulk$ regions are similar in both systems, 
as these are mostly outside of the bilayer and should not be affected by the new parameter set.

\subsection{Bilayer Electrostatics}

We further characterize the electrostatic properties of our bilayer
systems by computing the electrostatic potential across the simulation box. 
We do this following the protocol used in Saunders \etal
~\cite{saunders:2019}. We first compute the charge density of the
system components. We integrate this distribution twice, setting both
constants of integration to be zero to enforce a zero value for the
electric field in bulk solvent and a zero electrostatic potential at
the box edge. This is accomplished by taking the average value of the
electric field in the \emph{bulk} region of the box defined earlier,
and subtracting this value from all points. Due to the larger system
size in the optimized system, we needed to compute the average value of a much
larger region than in LB in order to apply boundary conditions. We
then integrate again to get the electrostatic potential. This result
can be seen in figure~\ref{fig:ch1}.
\begin{figure}[H]
    \caption[Electrostatic potential]{ Electrostatic potential as a function of distance from bilayer center. The optimized
cross terms yield a small change in the location of the peak of the potential in the
bilayer simulated with optimized cross terms, as well as the loss of the valley behind the peak.
}
    \label{fig:potential:ch1}
    \includegraphics[width=0.5\textwidth,trim=-0cm 0 0 0]{figure_s3_ch1.eps}
\end{figure}
The shape of the potential
is largely unaltered within fluctuations. Systems simulated with the
optimized parameters and with LB rules both have a similar bilayer dipole potential, 
which remains elevated over
the system without salt, by $\sim$ 220~mV.  We report 
that the optimized system has a slightly
elevated bilayer dipole potential compared to the system simulated with
LB rules, increased by $\sim$ 12~mV.
This may be a direct result of the larger number of
ions bound to the bilayer in this system.  We also note the system simulated with optimized cross terms
has different details
throughout the electrostatic potential compared to the system simulated with LB rules and in the
system without salt, however these are within fluctuations and cannot be used to draw conclusions.

Poisson--Boltzmann (PB) theory is a mean field approximation for
solvated ions near an interface. Experimentally PB theory is used to
assess the surface potential of the lipid bilayers. We also examine
the behavior of the ions in bulk solvent under the framework of PB
theory. Following the procedure used in our previous
work~\cite{saunders:2019}, we fit the number density of \cl ions in
the solvent--occupied region of the box to a Poisson-Boltzmann
distribution, using the inverse Debye length $K$ and the density of
\cl at the center of the solvent occupied region of the box $\rho_0$
as fit parameters. 
The density is modeled as:
\begin{equation}
    \label{eq:gcdens}
    \rho (z)= \rho _{0} \exp{(- \bar{z} e \beta \psi (z))},
\end{equation}
where $\rho_0$ is the number density of the ion at the center of the
solvent--occupied region of the box, $\bar{z}=1$ is the valency of the
ion in the system, $\beta=\frac{1}{k_b T}$, $e$ is the charge on an
electron, and $\psi(z)$ is the electrostatic potential. We then assume
the form of $\psi(z)$ to be the sum of two Debye-Huckel
potentials~\cite{israelachvili:2011:intermol} reflected across the
center of the solvent--occupied region of the box:
\begin{equation}
    \begin{split}
    \psi_1(z)&=\psi_s \exp(-K(z+\frac{D}{2})) \\
    \psi_2(z)&=\psi_s \exp(K(z+\frac{D}{2})) \\
    \psi(z)&=\psi_1 + \psi_2\text{,}
    \label{eq:gcpot}
\end{split}
\end{equation}
where $D$ is the distance from the hydration boundary of one bilayer leaflet
to the next across the solvent, $K$ is the inverse Debye length, and $\psi_s$ is the surface potential:
\begin{equation}
    \psi_s = \frac{\varsigma}{\varepsilon_0\varepsilon K}\text{.}
\label{eq:gcspot}
\end{equation}
%, given by
%\begin{equation}
%K=\sqrt{\sum_i{\rho_{0,i}\bar{z^2_i}\frac{e^2}{\epsilon_0 \epsilon k_bT}}}
%\label{eq:gcK}
%\end{equation}
\deleted{We used a value of $D=13.167$ nm for the LB system, and $D=27.01$ nm for the system simulated with optimized parameters. }
\added{The LB system yielded a value of $D=13.167$ nm and the system }
\added{simulated with optimized parameters, containing twice as many }
\added{solvent molecules, }
\added{gave a value of $D=27.01$ nm.}
We take the surface charge density $\varsigma$ from the charge density inside
of the hydration boundary of the lipid bilayer. Since only ions
contribute a net charge to our system, we compute this using only the
charge density of ions in the system. 
This value was computed to be $\varsigma=0.13$ \emph{e} nm$^{-2}$ for the system simulated with LB rules, and 
$\varsigma=0.11$ \emph{e} nm$^{-2}$ for the system simulated with the new parameters.
Our fitting procedure yielded 
number densities $\rho _{0}=0.043\text{nm}^{-3}$ for the system simulated with LB rules, 
and $\rho _{0}=0.079\text{nm}^{-3}$ for the system simulated with optimized parameters.
The fitted inverse screening lengths were found to be $K=0.91 \pm 0.014$ nm$^{-1}$ for the LB rules simulation 
and $0.94 \pm 0.018$ nm$^{-1}$ for the system simulated with optimized parameters.
The resulting fit and predicted
density of \na ions and electrostatic potential can be seen in
figure~\ref{fig:gouy:ch1}. 
\begin{figure}[h!]
    \caption[Poisson-Boltzmann theory]{ Poisson-Boltzmann theory predictions and simulation results. (a) shows a snapshot of the 
system simulated with optimized cross terms, translated
to center the solvent occupied region. Water has been hidden
for clarity. (b) and (d)
    show the number density of ions in the solvent occupied region of the box. 
    (c) and (e) show the corresponding electrostatic potential in solvent.
    We illustrate theoretical predictions as solid lines, with corresponding 
    simulation results as points with error bars. Red vertical lines denote the \emph{hydration boundary} of the lipid bilayer. 
    \cl density data is used for fitting in both systems. 
}
    \label{fig:gouy:ch1}
    \includegraphics[width=0.5\textwidth,trim=0 0 0 0]{figure_s4_ch1.eps}
\end{figure}
We see the results from our simulation
represented by points with error bars, while PB theory results are
shown in solid lines.  We see excellent agreement in the \na density
profile away from the bilayer surface, and reasonable agreement in the
electrostatic potential. From this we can see that the optimized and LB
systems both exhibit similar ionic distributions with models used to
describe electrophoretic mobility
experiments~\cite{israelachvili:2011:intermol}.

\section{Conclusions}

Mixing rules are often relied upon to compute non--bonded cross terms 
for interacting molecules in molecular simulations. 
However, when mixing force--fields that have been developed independently of each other, 
inaccuracies may develop. 
Here we demonstrate one such case and propose a rigorous solution. 
MD simulations conducted using predefined mixing rules for 
non-polarizable force fields developed separately for ions and lipids 
have always produced very pronounced salt-induced structural changes in lipid bilayers. 
Contrary to this, most experimental observations point to a 
moderate or even an insignificant change in bilayer structure at physiological salt concentrations. 
We resolve this discrepancy by explicitly parameterizing ion--lipid cross terms
using our procedure {\em ``Many Body Non Bonded fix''} (MB-NB-fix). 
It is based on the NB-fix method employed in previous works
~\cite{baker:2010:accurate,yoo:2012:improved,fyta:2012:ionic,mamatkulov:2013:force,venable:2013,savelyev:2014:balancing,li:2015:representation,savelyev:2015:competition,jing:2017:study}
and utilizes ParOpt software developed in our lab~\cite{fogarty:2014:paropt}\cite{fogarty:2014:thesis}.
We note that after applying the optimized parameters for \na--lipid interactions, the 
bilayer structure conforms more to experimental
observations while all other properties such as solvent structure,
electrostatic potential, and dynamic properties are approximately
similar to that obtained with those obtained with LB parameters. 
We note that we have not applied this method to optimize \cl~ interactions terms, which
may still further affect the bilayer structure. This will be the subject of future work.

The MB-NB-fix method proposed here is a 
general method which can be used to derive mixing
terms for simulations with independently developed force fields. 
This method will be used in future work to improve other sets of mixed force-fields, including 
those of other monovalent ions and the gromos 43A1-S3 lipids, and between these lipids and amino-acids for use in proteins.
Furthermore, many body cooperativity effects, such as ion-induced polarization 
in lipid molecules may be critical to further improving
the reproduction of lipid bilayer structure.
A correct approach to incorporate these effects to our simulation 
would be to have explicit polarization terms in 
our simulation models. This is complicated, as most existing polarizable 
simulation models are either not very effective
at accurately reproducing polarization effects or are
much more computationally expensive compared
to classical non-polarizable simulations. The MB-NB-fix method
has potential to become an ideal solution for mixing force-fields, including
polarizable and non-polarizable models in the same system 
to construct simulations that are tractable yet accurate.